---
title: "HW-Classification"
author: "Sena Çiftçi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

## Credit Data 

This assignment is related with Credit Data. As the first step, the provided dataset must be imported to R.  

```{r import}
# libraries to be used 
library(tidyverse)
library(GGally)
library(rpart)
library(rpart.plot)
library(cvms)
library(tibble)

# import the data set
credit <- read_csv(file="credit.csv")

# convert to data frame 
credit <- as.data.frame(credit)

dim(credit)
colnames(credit)
str(credit)

```


## Variables  

person_age: Age    
person_income: Annual Income    
person_home_ownership: Home ownership    
person_emp_length: Employment length (in years)     
loan_intent: Loan intent    
loan_grade: Loan grade    
loan_amnt: Loan amount    
loan_int_rate: Interest rate    
loan_status: Loan status (0 is non default 1 is default)    
loan_percent_income: Percent income    
default_before: Historical default     
cred_hist_length: Credit history length     


## First Look: Summary and Glimpse

A first look: 

- How many observations?  

- How many NAs?

- Categorical Variable Names?  

- Tables for Categorical Variables?  



```{r}

summary(credit)
glimpse(credit)

# How many observations?

nrow(credit)

# How many NAs?

colSums(is.na(credit))

# Numeric variables
credit |> 
  select_if(is.numeric) |> # select only numeric columns
  colnames()  # names 

# Character variables
credit |> 
  select_if(is.character) |> # select only character columns
  colnames() #names

# Analyze the class imbalance of the loan_status variable

table(credit$loan_status)

# Meaning the 22427 of the borrowers have successfully repaid the loan (low risk) and the 6203 of the borrowers have not successfully repaid the loan (high risk)

# table of home_ownership

table(credit$home_ownership)

# more tables
table(credit$loan_intent)
table(credit$loan_int_rate)

```

## Obtaining Subset of Data 

For this assignment we first obtain the subset of the dataset before EDA and modelling   


```{r}

set.seed(338011) # last 6 digits of my student id
cr_r <- sample(1:nrow(credit), size = 7500) # rows
cr_c <- c(9, 1:3, 7, 10 ,sample(c(4:6,8,11,12),2)) # columns 

credit_data <- credit[cr_r,cr_c] # sub-sample of data

credit_data |>
  select_if(is.numeric)
  ggpairs(credit_data)

glimpse(credit_data)

```

## Some Tables 

```{r}
credit_data |> 
  group_by(loan_status) |> 
  summarise(m_age = mean(age, na.rm=T),
            m_income = mean(income, na.rm=T),
            m_amount = mean(loan_amnt, na.rm=T),
            m_perc_income = mean(loan_percent_income, na.rm=T)
  )

table(credit_data$loan_status) # shows whether borrowers repaid successfully or not
summary(credit_data)



table(credit_data$home_ownership) # status for home ownership
table(credit_data$loan_intent)
table(credit_data$loan_grade) # statuses and numbers for loan grades


credit_data |>
  group_by(loan_status) |>
  count(home_ownership) # status of home ownership categorized by whether or not the borrowers repaid successfully or not

# Findings: Most borrowers have successfully repaid the loan
# While, in other statuses, the rate of borrowers who repaid the loan is about the double the size of borrowers who have not repaid successfully. But in 'mortgage' it is over 6 times and in  'own' it is over 13 times

```



## Let's Create [Some] Plot[s]   


```{r}
credit_data |> 
  ggplot(aes(loan_percent_income, loan_amnt)) + 
  geom_point(color="blue") + 
  facet_wrap(~ loan_status, 
             labeller = as_labeller(c("0" = "Not Default", "1"="Default"))) + 
  labs(
    title = "Loan Percent Income vs Loan Amount", 
    x = "Loan Percent Income", 
    y = "Loan Amount"
     ) + 
  theme_bw()

credit_data |>
  select_if(is.numeric) |>
  ggpairs()

# own trials (box plot of the sample)

credit_data |> 
  ggplot(aes(loan_percent_income, loan_amnt)) + 
  geom_boxplot() + 
  facet_wrap(~ loan_status, 
             labeller = as_labeller(c("0" = "Not Default", "1"="Default"))) + 
  labs(
    title = "Loan Percent Income vs Loan Amount",
    x = "Loan Percent Income", 
    y = "Loan Amount"
  ) + 
  theme_bw()

# add more as you need

credit_data |> 
  ggplot(aes(x = factor(loan_status), y = income)) +
  geom_point() +
  labs(
    x = "Loan Status",
    y = "Income"
  ) +
  theme_bw()

#box plot

credit_data |> 
  ggplot(aes(x = factor(loan_status), y = income)) +
  geom_boxplot() +
  labs(
    x = "Loan Status",
    y = "Income"
  ) +
  theme_bw()

```


## Linear Probability Model      

- First Linear Model: 

    + Dependent Var: loan_status
    + Independent Vars: Start with all predictors

Give a brief summary of linear model, What you see from the summary table (interpret estimations, coefficient significance etc)   

```{r}

lm_1 <- lm(loan_status ~ loan_percent_income + cred_hist_length
           + loan_amnt + age + income + home_ownership
           + loan_grade, data=credit_data)  

summary(lm_1)
```


- Create 2 more models containing variable transformation. Get the summary tables, interpret them  

- Select the best model among them, explain your reasoning   

```{r}

lm_2 <- lm(loan_status ~ log(income), data=credit_data)

names(lm_2)
summary(lm_2)


lm_3 <- lm(loan_status ~ income + I(income^2) + loan_amnt 
           + home_ownership, data=credit_data)  

names(lm_3)
summary(lm_3)

```


## Logistic Regression   

- First Logistic Model: 

    + Dependent Var: loan_status
    + Independent Vars: Start with all predictors   

Give a brief summary of linear model, What you see from the summary table (interpret estimations, coefficient significance etc)   

```{r}
# estimate a Logit regression with multiple regressors
logit_1 <- glm(loan_status ~ loan_percent_income + cred_hist_length
           + loan_amnt + age + income + home_ownership
           + loan_grade, 
                  family = binomial(link = "logit"), 
                  data = credit_data)

summary(logit_1)

```

## Decision Tree  

```{r}
tree_model <- rpart(
  loan_status ~ loan_percent_income + cred_hist_length
           + loan_amnt + age + income + home_ownership
           + loan_grade, 
  data = credit_data,
  method = "class"
)

rpart.plot(tree_model)

tree_model$variable.importance

```

## Summary

- Confusion Matrix for all models must be provided and compared with each other   


